---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(dplyr)
library(ggplot2)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data


# Measures of spread - variance, standard deviation and inter quartile range

Robust stastics - is the meadian or the mean a more robust statis

Median (Center) and IQR (Spread) are considered more robust than Mean and Standard Deviation because they are less seceptible to skew by extreme observations. Mean and SD are non-robust and are better suited to symetric distributions.

# Transformation of data

A transformation is a rescaling of data using a function. when data is strongly skewed it is sometimes better to transorm them so they are easier to model.

Common transformation - natural log. Opten applied when musch of the data cluster near zero relative to the larger values in the data.sqrt transformation to make results more linear.

Goals of transformation - to see the data structure differently. Reduce skew to assist in modeling. To straighten a nonlinear relationship in a scatterplot.
* * *

## Part 2: Research questions

**Research quesion 1:**

**Research quesion 2:**

**Research quesion 3:**


* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.


Create contingency tables!!
Relative frequencites - test categorical variables amongst stratafied groups (i.e. Income vs Difficulty saving) and determine if they are associated (dependent)



Categorical Variable:
Generate a mosaic plot and contingency table
Segmented bar plot to visualise relative frequenxies



Nul Hypotheses - nothing is going on (the status quo). Alternative hypothesis - there is something going on.
H0 - Defendant is innocent - HA Defendant is guilty.

Present the evidence, judge the evidence and either accept or reject the hypothesis.

**Research quesion 1:**

```{r}

```



**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```

