---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(dplyr)
library(ggplot2)
```

### Load data

```{r load-data}
load("brfss2013.RData")
```

* * *
## Part 1: Data - Method, Bias, Scope

The Behavioral Risk Factor Surveillance System (BRFSS) - 2013 is a surveillance system run by the Centers for Disease Control and Prevention that collects data from all US states and territoies by way of surveys. These surveys are run continuously through the year by contracted call centres, universities and in-house interviewers to provide detailed information regarding risk-behaviours, chronic health conditions and the use of preventive services. The dataset is drawn from over 400,000 adult interviews each year and is conducted via Random Digit Dialing to both landlines and cell phones.

Though the survey is adminstered throughout the year and at different times in the day there is inherent bias involved in collecting data via a phone sample. People that do not have a mobile or phone line, and people who don't pick up the phone are examples of people missed by the survey. Furthermore, the misremembering of activites, and the over or under exaggeration of particular traights may produce skewed results.

As an observational study with a randomised population the results are generalisable and corellation can be discussed, however as the subjects are not randomly assigened as in an experiment we cannot prove causality.

* * *

## Part 2: Research questions

# Come up with at least three research questions that you want to answer using these data. You should phrase your research questions in a way that matches up with the scope of inference your dataset allows for. Make sure that at least two of these questions involve at least three variables. You are welcomed to create new variables based on existing ones. With each question include a brief discussion (1-2 sentences) as to why this question is of interest to you and/or your audience.

**Research quesion 1:**
Is income level tied to mental health? 

Variable 1:income2 - categorical
Variable 2: menthlth - Finite Numerical 1-30 
Variable 3: ratesmenthealth - create new variable - menthlth <10 few, >=11 <=20 many, >=21 <=30 most 

**Research quesion 2:**
Do home owners vote more frequently than renters? Does gender change this?
Variable 1: scntvot1 - voting boolean
Variable 2: renthom1 - living arrangements catagorical
Variable 3: sex - categorical 

**Research quesion 3:**
Does a person's internet use correlate with a person's general health? Is there a difference in genders
Variable 1: internet (Use in the past 30 days) - Boolean
Variable 2: genhlth
Variable 3:

* * *

## Part 3: Exploratory data analysis

# Perform exploratory data analysis (EDA) that addresses each of the three research questions you outlined above. Your EDA should contain numerical summaries and visualizations. Each R output and plot should be accompanied by a brief interpretation. 

Create contingency tables!!
Relative frequencites - test categorical variables amongst stratafied groups (i.e. Income vs Difficulty saving) and determine if they are associated (dependent)

Categorical Variable:
Generate a mosaic plot and contingency table
Segmented bar plot to visualise relative frequenxies


**Research quesion 1:**

```{r}
q1 <- select(brfss2013, income2, menthlth) %>%
  filter(!is.na(income2), !is.na(menthlth))
q1 %>% group_by(income2) %>% summarise(count=n())
```



**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```

