---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(dplyr)
library(ggplot2)
```

### Load data

```{r load-data}
load("brfss2013.RData")
```

* * *
## Part 1: Data - Method, Bias, Scope

The Behavioral Risk Factor Surveillance System (BRFSS) - 2013 is a surveillance system run by the Centers for Disease Control and Prevention that collects data from all US states and territoies by way of surveys. These surveys are run continuously through the year by contracted call centres, universities and in-house interviewers to provide detailed information regarding risk-behaviours, chronic health conditions and the use of preventive services. The dataset is drawn from over 400,000 adult interviews each year and is conducted via Random Digit Dialing to both landlines and cell phones.

Though the survey is adminstered throughout the year and at different times in the day there is inherent bias involved in collecting data via a phone sample. People that do not have a mobile or phone line, and people who don't pick up the phone are examples of people missed by the survey. Furthermore, the misremembering of activites, and the over or under exaggeration of particular traights may produce skewed results.

As an observational study with a randomised population the results are generalisable and corellation can be discussed, however as the subjects are not randomly assigened as in an experiment we cannot prove causality.

* * *

## Part 2: Research questions

**Research quesion 1:**
Is income level tied to mental health? It would be interesting to see if there is an association present between mental health and income. Can money buy happiness?
<ul>
<li>Variable 1:income2 - categorical</li>
<li>Variable 2: menthlth - Finite Numerical 1-30</li> 
<li>Variable 3: ratesmenthealth - create new variable - menthlth <10 few, >=11 <=20 many, >=21 <=30 most</li> 
</ul>
**Research quesion 2:**
Do home owners vote more frequently than renters? Does gender change this? Home ownership typically resides with older people and is a proxy for young vs old.
<ul>
<li>Variable 1: scntvot1 - voting boolean</li>
<li>Variable 2: renthom1 - living arrangements catagorical</li>
<li>Variable 3: sex - categorical </li>
</ul>
**Research quesion 3:**
Does a person's internet use correlate with a person's general health? Is there a difference in genders? I would believe that the more time someone spends on the internet precludes them from living an active lifestyle which would indicate lower healthy outcomes.
<ul>
<li>Variable 1: internet (Use in the past 30 days) - Boolean</li>
<li>Variable 2: genhlth</li>
<li>Variable 3:</li>
</ul>
* * *

## Part 3: Exploratory data analysis

Perform exploratory data analysis (EDA) that addresses each of the three research questions you outlined above. Your EDA should contain numerical summaries and visualizations. Each R output and plot should be accompanied by a brief interpretation. 

**Research quesion 1:**

```{r}
# Select relevant variables, filter NAs, create a new categorical variable and summarise. 
q1 <- select(brfss2013, income2, menthlth) %>% 
  filter(!is.na(income2), !is.na(menthlth)) %>%
  mutate(menthlthcat = ifelse(menthlth > 15, c("High"), c("Low")))

q1 %>%
  group_by(income2, menthlthcat) %>% 
  summarise(count=n())

ggplot(data = q1, aes(x = menthlthcat, y = income2)) +
  geom_bar(stat = "count")
```


ggplot(data = q1, aes(x = menthlthcat) +
  geom_histogram())
  


**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```

